Bounds on Coding Theory from Algebraic Geometry

### 2021-04-06 draft.tex
Updating draft

\theoremstyle{theorem}
\newtheorem{thm}{Theorem}[section]

\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{alg}[thm]{Algorithm}

@@

\theoremstyle{example}
\newtheorem{prob}[thm]{Problem}

@@

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\Ima}{Im}
\DeclareMathOperator{\dimension}{dim}

@@

\maketitle
\begin{abstract} % NOT REQUIRED
    Coding theory is concerned with finding efficient ways to encode a message so that one may correct errors in the message. In algebraic coding theory, we study efficient codes generated from algebraic geometric methods.

    In this paper, I will construct the Reed--Solomon codes, generalize them using projective curves, and understand the results from \cite{TVZ82} on finding a bound better than the well-known Gilbert--Varshamov one.
\end{abstract}

@@

@@

\begin{defn}[Reed--Solomon Codes]
	Let $q$ be a power of a prime, and $\mathbb{F}_q = \{\alpha_1, \alpha_2, \dots, \alpha_q\}$ the field with $q$ elements. Let $k$ be an integer, and $\mathcal{L}_k$ the set of all polylomials over $\mathbb{F}_q$ with degree smaller than $k$. Let $k \le n \le q$ be an integer. The Reed--Solomon code $RS_q(n, k)$ over $\mathbb{F}_q$ is 
	\begin{equation}
		RS_q(n, k) = \left\{\left(f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)\right) \mid f \in \mathcal{L}_k\right\}.
	\end{equation}
\end{defn}

@@

\begin{proof}
	$R_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $\mathcal L_k$ is a vector space over $\mathbb{F}_q$. Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$. Consider the map $\phi: \mathcal L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\end{equation}

	Note that the map $\phi$ is a linear transformation. Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polylomial, which implies $f = g$. Therefore, $\phi$ is also injective. This implies that the dimension of the domain $\mathcal L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in \mathcal L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots. Futhermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots. Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f, g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with \nameref{thm:singleton_bound} implies 
	\begin{equation}
		k + d_\text{min} = n + 1.
	\end{equation}

@@

The Reed--Solomon codes are good codes in the sense that key have the largest possible sum $k + d$ for their length $n$. However, Reed--Solomon codes are limited becuase their length is at most the the alphabet size.

\section{Generalized Reed--Solomon codes} \label{s:grs}
We shall redefine the Reed--Solomon codes using language related to a projective line. There is a way to replace the ``projective line'' with a ``projective plane curve'' and create other codes, called \emph{Generalized Reed--Solomon codes} or simply \emph{algebraic geometric codes}. We want large $R$ and $\delta$, and these codes yield 
	\begin{equation}
		\label{eq:g/n} R + \delta \ge 1 + 1/n - g/n,
	\end{equation}
	where $n$ is the number of rational points of a curve $X$, with genus $g$.

>>>>

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{alg}[thm]{Algorithm}

@@

\theoremstyle{example}
\newtheorem{prob}[thm]{Problem}
\newtheorem{que}[thm]{Question}

@@

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\Ima}{Im}
\DeclareMathOperator{\dimension}{dim}
\DeclareMathOperator{\Supp}{Supp}

@@

%\begin{abstract} % NOT REQUIRED
%    Coding theory is concerned with finding efficient ways to encode a message so that one may correct errors in the message. In algebraic coding theory, we study efficient codes generated from algebraic geometric methods.
%    
%    In this paper, I will construct the Reed--Solomon codes, generalize them using projective curves, and understand the results from \cite{TVZ82} on finding a bound better than the well-known Gilbert--Varshamov one.
%\end{abstract}

\maketitle

@@

WRITE INTRODUCTION WITH ALICE AND BOB.

@@

\begin{defn}[Reed--Solomon Codes]\label{defn:rs_codes}
	Let $q$ be a power of a prime, and $\mathbb{F}_q = \{\alpha_1, \alpha_2, \dots, \alpha_q\}$ the field with $q$ elements. Let $k$ be an integer, and ${L}_k$ the set of all polynomials over $\mathbb{F}_q$ with degree smaller than $k$. Let $k \le n \le q$ be an integer. The Reed--Solomon code $RS_q(n, k)$ over $\mathbb{F}_q$ is 
	\begin{equation}
		RS_q(n, k) = \left\{\left(f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)\right) \mid f \in {L}_k\right\}.
	\end{equation}
\end{defn}

@@

\begin{proof}
	$R_q(n, k)$ is a subset of $\mathbb{F}_q^n$, thus it has length $n$.  
	Note that $ L_k$ is a vector space over $\mathbb{F}_q$. Note that $\{1, x, x^2, \dots, x^{k-1}\}$ is a choice of basis for this vector space, thus it has dimension $k$. Consider the map $\phi:  L_k \to \mathbb{F}_q^n$ given by 
	\begin{equation}
		f \mapsto (f(\alpha_1), f(\alpha_2), \dots, f(\alpha_n)).
	\end{equation}

	Note that the map $\phi$ is a linear transformation. Thus, its image $\Ima \phi = RS_q(n, k)$ is also a vector space.  
	Additionally, if $\phi(f) = \phi(g)$, then $f - g$ has at least $n$ roots, but has degree less than $n$; thus $f - g$ is the zero polynomial, which implies $f = g$. Therefore, $\phi$ is also injective. This implies that the dimension of the domain $ L_k$ is the same as the dimension of the image $RS_q(n, k)$, i.e.,  $\dimension RS_q(n, k) = k$.

	Finally, consider distinct $f, g \in  L_k$ and define $d = \dist(\phi(f), \phi(g))$, $f - g$ has at least $n - d$ roots. Furthermore, $f - g$ is a non-zero polynomial with degree less than $k$, thus has at most $k - 1$ roots. Then,  $ k - 1 \ge \#\ \text{roots} \ge n - d$.
	If we choose $f, g$ such that $d$ is the minimal distance, we get $k + d_{\mathrm{min}} \ge n + 1$, which together with \nameref{thm:singleton_bound} implies 
	\begin{equation}
		k + d_\text{min} = n + 1.
	\end{equation}

@@

The Reed--Solomon codes are very good codes in the sense that key have the largest possible sum $k + d$ for their length $n$. However, Reed--Solomon codes are limited because their length is at most the alphabet size. So, a question naturally arises: Given fixed $\mathbb{F}_q$, are there codes over $\mathbb{F}_q$ with arbitrarily large $n$ and $R + \delta = 1 + 1/n$? If not, how large can $R$ and $\delta$ be when $n$ gets larger? The Gilbert--Varshamov bound shows that there are codes with 
\begin{equation}
	1 - R \approx q(\delta), \text{as $n \to \infty$,}
\end{equation}
in which
\begin{equation}
	q(x) = x \log_q(q-1) - x\log_q(x) - (1-x)\log_q(x).
\end{equation}
Are there any better codes?

\section{Rational Functions and Divisors}

\begin{defn}[Rational Function]
	A rational function $f$	is a function which is the ratio $g/h$ of two polynomials.
	It is homogeneous if $g, h$ are homogeneous.
	After cancelling common roots of $g, h$, the roots of $g$ are called \emph{zeros} of $f$ and the roots of $h$ are called the \emph{poles} of $f$.
\end{defn}

	We say $f$ has order $n$ in $P$ if $P$ is a zero of muliplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; order $0$, otherwise.

\begin{defn}[Divisor] 
	Let $F$ be the algebraic closure of $\mathbb{F}_q$.
	Let $X$ be an irreducible nonsingular projective curve in $N$-dimensional projective space over $F$.
	A \emph{divisor} on $X$ is a formal finite sum of the form $D = \sum a_P P$, where $P$ are points of $X$, $a_P$ are integers and $a_P = 0$ for all but finitely many points $P$.
	The \emph{degree} of $D$ is $\sum n_P$.
	The \emph{support} $\Supp D$ is the set $\{P \in X : a_P \neq 0\}$
\end{defn}

If $D = \sum n_p P$, then define the vector space $\mathcal L(D)$ as the set of all homogeneous rational functions $f$ such that the order of $f$ at each point $P$ of $X$ is greater or equal to $n_P$.
For our study, an important theorem is the following:
\begin{thm}[Riemman--Roch Theorem, \cite{Wal00}]
	Let $X$ be a nonsingular projective curve of genus\footnote{For the reader that is not familiar with genus, it is enough to know that it is an integer that can be calculated for any given curve.} $g$ defined over the field $\mathbb{F}_q$ and let $D$ be a divisor on $X$. Then
	\begin{equation}
		\dim \mathcal L(D) \ge \deg D + 1 - g,
	\end{equation}
	with equality holding if $\deg D > 2g - 2$.
\end{thm}

%Note that, if the degree of $D$ is negative, then
%\begin{equation}
%	\mathcal L(D) = \{0\}.
%\end{equation}
%Additionally, the \emph{Riemann--Rock theorem} implies that the dimension of the vector space $\mathcal L(D)$ is 
%\begin{equation}
%	 \dim\mathcal L(D) \ge \deg D - g + 1,
%\end{equation}
%where $g$ denotes the genus of $X$ (for the reader who is not familiar with the term genus, it suffices to know that it is an integer which can be calculated for any given curve).

\section{Generalized Reed--Solomon codes} \label{s:grs}
Let $\mathbb{P}^1(\mathbb{F}_q)$ denote the projective line over $\mathbb{F}$. We will write $(a : b)$ to denote the projective point corresponding to the $1$-dimensional vector space through $(a, b)$. The points on $\mathbb{P}^1(\mathbb{F}_q)$ are the points
\begin{equation}
	P_i = (\alpha_i : 1), \qquad 1 \le i \le q,
\end{equation}
and
\begin{equation}
	P_\infty = (1 : 0).
\end{equation}

Following \cite{LS87}, let $\mathcal L_k$ be the set of two-variable homogeneous rational functions which have a pole of order less than $k$ in the point $Q$.

\begin{prop}
	The sets $\mathcal L_k$ and $L_k$ are mapped with a bijection $\phi: f(x) \mapsto f(x/y)$.
\end{prop}
\begin{proof}
	WRITE PROOF.
\end{proof}

Then, we can rewrite the Reed--Solomon code from \ref{defn:rs_codes} as
\begin{equation}
	RS_q(n, k) = \{f(P_1), f(P_2), \dots, f(P_n) \mid f \in \mathcal L_k\}.
\end{equation}

\begin{defn}[Algebraic Geometric Code]
\end{defn}

We want large $R$ and $\delta$, and these codes yield 
	\begin{equation}
		\label{eq:g/n} R + \delta \ge 1 + 1/n - g/n,
	\end{equation}
where $n$ is the number of rational points of a curve $X$, with genus $g$.
